---
title: "Week Three Assignment - Data Preparation"
author: "Henry Lin"
output:
  html_document:
    theme: sandstone
    highlight: tango
    toc: yes
    toc_float: yes
    css: ../css/sandstone.css
  pdf_document: default
---


```{r, echo=F}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE) #Chunk default settings 
```

## Project Set-up
Library necessary packages 
```{r}
library(dplyr)
library(janitor)
```

## Load in original data 
Load in your raw data file - with `load` command for the `addhealth4` data
```{r}
load("/cloud/project/data/addhealth4.RData")
```


## New data frame with selected variables 
```{r}
mydata <- addhealth4 %>% 
  select(H4LM20, H4LM17, H4LM26, H4LM14, H4LM28, H4LM30, H4LM19, H4LM12)
```


## Recoding of variables

### Variable 1 (Time of Day Worked)
```{r}
mydata <- mydata %>% 
  mutate(time_of_day_worked = recode_factor(H4LM20,
                                "1"="regular day shift",
                                "2"="regular evening shift",
                                "3"="regular night shift",
                                "4"="shift rotates that is, it(changes/changed) periodically from day to evening or night",
                                "5"="split shift that is, it (consists/consisted) of two distinct periods each day",
                                "6"="irregular schedule or hours",
                                "7"="other",
                                "96"="refused",
                                "97"="legitimate skip",
                                .ordered=T))
```

### Variable 2 (Sepaeration Cause)
```{r}
mydata <- mydata %>% 
  mutate(seperation_cause = recode_factor(H4LM17,
                                "1"="layoff",
                                "2"="plant closed",
                                "3"="end of temporary or seasonal job",
                                "4"="discharged or fired",
                                "5"="program ended",
                                "6"="health problems",
                                "7"="pregnancy or family reasons",
                                "8"="return to school",
                                "9"="quit to look for another job",
                                "10"="quit to take another job",
                                "11"="retired",
                                "12"="military service",
                                "13"="incarceration",
                                "14"="other",
                                "96"="refused",
                                "97"="legitimate skip",
                                "98"="don't know",
                                .ordered=T))
```

### Variable 3 (Job Satisfaction)
```{r}
mydata <- mydata %>%
  mutate(job_satisfaction = recode_factor(H4LM26,
                                "1"="extremely satisfied",
                                "2"="satisfied",
                                "3"="neither satisfied nor dissatisfied",
                                "4"="dissatisfied",
                                "5"="extremely dissatisfied",
                                "96"="refused",  
                                "97"="legitimate skip",
                                "98"="don't know",
                                .ordered=T))                                          
```

### Variable 4 (What People Are Doing)
```{r}
mydata <- mydata %>%
  mutate(current_status = recode_factor(H4LM14,
                                "1"="only temporarily laid off",
                                "2"="on sick leave or temporarily disabled",
                                "3"="on maternity/paternity leave",
                                "4"="permanently disabled",
                                "5"="unemployed and looking for work",
                                "6"="unemployed and looking for work",
                                "7"="student",
                                "8"="keeping house",
                                "9"="retired",
                                "10"="other",
                                "96"="refused",  
                                "97"="legitimate skip",
                                .ordered=T))    
```

### Variable 5 (How Family Impacts Work)
```{r}
mydata <- mydata %>% 
  mutate(family_impact_on_work = recode_factor(H4LM28,
                                "1"="strongly agree",
                                "2"="agree",
                                "3"="neither agree or disagree",
                                "4"="disagree",
                                "5"="strongly disagree",
                                "96"="refused",
                                "98"="don't know",
                                .ordered=T))
```

### Variable 6 (How Work Impacts Time With Family)
```{r}
mydata <- mydata %>%
  mutate(work_impact_family_time = recode_factor(H4LM30,
                                                 "1"="frequently",
                                                 "2"="sometimes",
                                                 "3"="rarely",
                                                 "4"="never",
                                                 "6"="refused",
                                                 "7"="legitimate skip",
                                                 "8"="don't know",
                                                 .ordered=T))
```

### Variable 7 (Amount of Time Worked)

No need to recode.

### Variable 8 (Number of Jobs)
```{r}
mydata <- mydata %>%
  mutate(number_of_jobs_worked = recode_factor(H4LM12,
                                     "1"="1 job",
                                     "2"="2 jobs",
                                     "3"="3 jobs",
                                     "4"="4 jobs",
                                     "5"="5 jobs",
                                     "6"="6 jobs",
                                     "7"="7 jobs",
                                     "12"="12 jobs",
                                     "97"="legitimate skip",
                                     "98"="don't know",
                                     .ordered = T))
```

```{r}
mydata %>%
  tabyl(time_of_day_worked) %>%
  adorn_pct_formatting(digits=2)
```

```{r}
mydata %>%
  tabyl(job_satisfaction) %>%
  adorn_pct_formatting(digits = 2)
```

```{r}
mydata %>%
  tabyl(current_status) %>%
  adorn_pct_formatting(digits = 2)
```

```{r}
mydata %>%
  tabyl(family_impact_on_work) %>%
  adorn_pct_formatting(digits = 2)
```

```{r}
mydata %>%
  tabyl(work_impact_family_time) %>%
  adorn_pct_formatting(digits=2)
```

```{r}
mydata %>%
  tabyl(number_of_jobs_worked) %>% 
  adorn_pct_formatting(digits=2)
```

## Cleaned data frame

**Step 1** : Select only the variables that you want to keep, so you only have the version of the variables that you want in your data frame.
```{r}
clean <- addhealth4 %>% select(H4LM20, H4LM17, H4LM26, H4LM14, H4LM28, H4LM30, H4LM19, H4LM12)
```

**Step 2**: Save the resulting data set to your `data` folder as `datasetname_clean.Rdata` e.g. `addhealth_clean.Rdata`. 
```{r}
save(clean, file="/cloud/project/data/clean.Rdata")
save(mydata, file="/cloud/project/data/mydata.Rdata")
```

```{r}
glimpse(mydata)
glimpse(clean)
```

## Table of variables

Original Variable (from addhealth)  |   New Variable  |    Description (from addhealth)
-------------------|-----------------|------------------
H4LM20 | time_of_day_worked | Which one of these categories best describes the hours you (work/worked) at this job?
H4LM17 | separation_cause | What is the main reason you left your most recent job?
H4LM26 | job_satisfaction | How satisfied (are/were) you with this job?
H4LM14 | current_status | Which one of the following categories best describes what you're doing now?
H4LM28 | family_impact_on_work | Indicate how much you would agree or disagree with this statement: Family responsibilities have interfered with my ability to work.
H4LM30 | work_impact_family_time | (In the past 12 months/Since you started your current job/In the last year of your most recent job), how often on your primary job (have you spent/have you spent/did you spend) less time with your family than you wanted to because of work responsibilities?
H4LM19 | time_worked | How many hours a week (do/did) you usually work at this job? NOTE: Smallest 5 and largest 5 values are displayed.
H4LM12 | number_of_jobs_worked | On how many jobs are you currently working for pay at least 10 hours a week?