---
title: "Bivariate Graphing Assignment"
author: "Henry Lin"
output:
  pdf_document: default
  html_document: default
---

```{r, message=FALSE, warning=FALSE}
library(ggplot2)
library(knitr)
library(dplyr)
library(ggplot2)
library(janitor)
library(purrr)
library(kableExtra)
library(ggmosaic) 
library(scales)
opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
load("/cloud/project/data/clean.Rdata")
load("/cloud/project/data/current_status.Rdata")
load("/cloud/project/data/family_impact_on_work.Rdata")
load("/cloud/project/data/hours_worked.Rdata")
load("/cloud/project/data/job_satisfaction.Rdata")
load("/cloud/project/data/mydata.Rdata")
load("/cloud/project/data/number_of_jobs.Rdata")
load("/cloud/project/data/seperation_cause.Rdata")
load("/cloud/project/data/time_of_day_worked.Rdata")
load("/cloud/project/data/work_impact_family_time.Rdata")
load("/cloud/project/data/addhealth4.RData")
```

# Variables of Interest

Original Variable (from addhealth)  |   New Variable  |    Categorical or Numerical  |  Description (from addhealth)    
-------------------|-----------------|------------------|------------------
H4LM20 | time_of_day_worked | Categorical | Which one of these categories best describes the hours you (work/worked) at this job?
H4LM17 | separation_cause | Categorical | What is the main reason you left your most recent job?
H4LM26 | job_satisfaction | Categorical | How satisfied (are/were) you with this job?
H4LM14 | current_status | Categorical | Which one of the following categories best describes what you're doing now?
H4LM28 | family_impact_on_work | Categorical |Indicate how much you would agree or disagree with this statement: Family responsibilities have interfered with my ability to work.
H4LM30 | work_impact_family_time | Categorical | (In the past 12 months/Since you started your current job/In the last year of your most recent job), how often on your primary job (have you spent/have you spent/did you spend) less time with your family than you wanted to because of work responsibilities?
H4LM19 | hours_worked | Numerical | How many hours a week (do/did) you usually work at this job?
H4LM12 | number_of_jobs_worked | Numerical | On how many jobs are you currently working for pay at least 10 hours a week?

# C ~ C Association
time_of_day_worked is the shifts worked by the respondents. 
job_satisfaction is respondent's satisfaction with their employment.

```{r time_of_day/job_satisfaction table}
mydata%>%
  
  filter(job_satisfaction %in% c ("extremely satisfied", "satisfied", "neither satisfied nor dissatisfied", "dissatisfied", "extremely dissatisfied")) %>%

  droplevels() %>% 
  
  tabyl( bio_sex, job_satisfaction) %>%
    adorn_totals(c("row")) %>% 
    adorn_percentages("row") %>% 
    adorn_pct_formatting(digits=2) %>% 
  
  kable() %>%
    kable_styling(bootstrap_options = "striped",
                  full_width = T, 
                  position="left") %>%
  add_header_above(c("","job satisfaction"=5))
```

```{r time_of_day/job_satisfaction histogram}
mydata%>%
  
  filter(job_satisfaction %in% c ("extremely satisfied", "satisfied", "neither satisfied nor dissatisfied", "dissatisfied", "extremely dissatisfied")) %>%
  droplevels() %>% 
  
  ggplot(aes(x=bio_sex, fill=job_satisfaction))+
  geom_bar(position = position_fill(reverse = T), width=.6) +
  
  scale_fill_brewer(palette="Dark2") +
  scale_y_continuous(labels=percent) +
  coord_flip()
```

## Narrative
When looking at the all addhealth4observations, the rates of job satisfaction are similar for both sexes, showing no potential association. When comparing job satisfaction with gender, 23.47% of men responsed with extremely satisfied, compared to 23.80% for women, for difference in sample proportions of 0.33%. 
49.91% of men responsed with satisfied, compared to 49.82% for women, for difference in sample proportions of 0.09%.
18.02% of men responsed with neither satisfied nor dissatisfied, compared to 17.23% for women, for difference in sample proportions of 0.79%. 
6.14% of men responsed with dissatisfied, compared to 6.90% for women, for difference in sample proportions of 0.76%
2.46% of men responsed with extremely dissatisfied, compared to 2.25% for women, for difference in sample proportions of 0.21%.  
The relationship between sex and job satisfaction was similar for men and women.

# Q ~ C Association

```{r job_satisfaction/hours_worked table}
mydata %>% 
  
  filter(hours_worked < 121) %>% 
  
  group_by(bio_sex) %>%
  
  summarize(n = n(),
            min = min(hours_worked),
            q1 = quantile(hours_worked, probs=0.25),
            median = median(hours_worked),
            mean = round(mean(hours_worked),1),
            q3 = quantile(hours_worked, probs=0.75),
            max = max(hours_worked),
            sd = round(sd(hours_worked),1)) %>% 
  
  kable() %>% 
    kable_styling(bootstrap_options = "striped",
                  full_width = T, 
                  position="left")   %>% 
    add_header_above(c("hours worked "= 9))
```

```{r job_satisfaction/hours_worked histogram}

mydata %>% 
  
  filter(hours_worked < 121 ) %>%
  
  ggplot(aes(hours_worked))+
  
    geom_histogram(aes(y=..density..,
                       fill=bio_sex), 
                   color="white", 
                   binwidth = 5) +
  
      scale_fill_brewer(palette="Dark2") +
  
    facet_wrap(~ bio_sex, 
               ncol=1,
               as.table = FALSE)  ## reverses the order of the levels to match boxplots
```

> Calculate grouped summary statistics (means, sd by group) and compare them using sentences with numbers to justify your claims. Compare center, shape and spread of the quantitative variable across different levels of the categorical variable. 

## Narrative
Although both males and females work a median of 40 hours and worked the same minimum and maximum hours, 10 hours and 120 hours respectively, males work a mean of 44 hours, compared to women, at 38.8 hours. This represents a difference of 13.40% of working hours for males compared to females. There are further differences in addition. 
The interquartile range for males is higher at 40 to 50 hours, compared to women at 35 to 40 hours. Meaning that the 

# (B ~ C, C ~ B, or Q ~ B)
> same format as above


# Q ~ Q Association
> Calculate the correlation and use it when discussing the direction, strength, and form of this relationship. 
