---
title: "Bivariate Graphing Assignment"
author: "Henry Lin"
output:
  pdf_document: default
  html_document: default
---

```{r, message=FALSE, warning=FALSE}
library(ggplot2)
library(knitr)
opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)

load("/cloud/project/data/clean.Rdata")
load("/cloud/project/data/current_status.Rdata")
load("/cloud/project/data/family_impact_on_work.Rdata")
load("/cloud/project/data/hours_worked.Rdata")
load("/cloud/project/data/job_satisfaction.Rdata")
load("/cloud/project/data/mydata.Rdata")
load("/cloud/project/data/number_of_jobs.Rdata")
load("/cloud/project/data/seperation_cause.Rdata")
load("/cloud/project/data/time_of_day_worked.Rdata")
load("/cloud/project/data/work_impact_family_time.Rdata")
load("/cloud/project/data/addhealth4.RData")
```

# C ~ C Association
> Describe each variable in words (what does it say in the codebook) then delete this text. 

Original Variable (from addhealth)  |   Modified Variable Name  |    Categorical or Numerical  |  Description (from addhealth)    
-------------------|-----------------|------------------|------------------
H4LM20 | time_of_day_worked | Categorical | Which one of these categories best describes the hours you (work/worked) at this job?
H4LM17 | separation_cause | Categorical | What is the main reason you left your most recent job?
H4LM26 | job_satisfaction | Categorical | How satisfied (are/were) you with this job?
H4LM14 | current_status | Categorical | Which one of the following categories best describes what you're doing now after leaving your last job?
H4LM28 | family_impact_on_work | Categorical |Indicate how much you would agree or disagree with this statement: Family responsibilities have interfered with my ability to work.
H4LM30 | work_impact_family_time | Categorical | (In the past 12 months/Since you started your current job/In the last year of your most recent job), how often on your primary job (have you spent/have you spent/did you spend) less time with your family than you wanted to because of work responsibilities?
H4LM19 | time_worked | Numerical | How many hours a week (do/did) you usually work at this job?
H4LM12 | number_of_jobs_worked | Numerical | On how many jobs are you currently working for pay at least 10 hours a week?

```{r}
# put your code to create the plot here
```

> Describe the association between the two variables here in paragraph form. Use N and %'s in your answer. Then delete this text. 


# Q ~ C Association
> same format as above
> Calculate grouped summary statistics (means, sd by group) and compare them using sentences with numbers to justify your claims. Compare center, shape and spread of the quantitative variable across different levels of the categorical variable. 


# (B ~ C, C ~ B, or Q ~ B)
> same format as above


# Q ~ Q Association
> Calculate the correlation and use it when discussing the direction, strength, and form of this relationship. 
